---
title: "Cross-cohort PDAC transcriptomic atlas"
author: "Analysis pipeline"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  root.dir = normalizePath(getwd(), winslash = "/", mustWork = TRUE)
)
library(tidyverse)
read_tsv_safe <- function(path) {
  if (file.exists(path)) readr::read_tsv(path, show_col_types = FALSE) else tibble(note = paste("Missing file:", path))
}
show_head <- function(path, n = 30) {
  x <- read_tsv_safe(path)
  knitr::kable(head(x, n))
}
```

# Dataset and metadata audit

Manual GEO audit confirms that `GSE71989` contains 13 PDAC/tumor samples and 8 control samples. For `GSE91035`, the confirmed labels are 25 tumor/PDAC, 8 control, 15 benign, and two chronic-pancreatitis samples (`GSM2420007`, `GSM2420010`). CP samples are retained in the metadata audit as `cp` but are excluded by default from the primary tumor/control and normal-benign-PDAC transition analyses.

## Dataset sample summary

```{r}
show_head("results/tables/dataset_sample_summary.tsv", 100)
```

## Expected versus observed biological labels

```{r}
show_head("results/tables/metadata_expected_vs_observed.tsv", 100)
```

## Samples requiring manual review

```{r}
show_head("results/tables/metadata_samples_requiring_review.tsv", 100)
```

# Original atlas programs

## Core PDAC tumor-associated program

```{r}
show_head("results/tables/program_core_tumor_signature.tsv", 30)
```

## Microenvironment-sensitive program

```{r}
show_head("results/tables/program_microenvironment_signature.tsv", 30)
```

## Normal-benign-PDAC transition program

```{r}
show_head("results/tables/program_benign_to_pdac_transition.tsv", 30)
```

# Manuscript-reinforced programs

## Reinforcement summary

```{r}
show_head("results/tables/reinforcement_summary.tsv", 50)
```

## Tiered cross-cohort signature

Main-text core claims should rely on Tier 1 genes. Tier 2 and platform-limited genes are retained for supplementary or hypothesis-generating analyses.

```{r}
show_head("results/tables/tiered_core_signature.tsv", 40)
```

## Tier 1 main-text core signature

```{r}
show_head("results/tables/program_tier1_core_signature.tsv", 40)
```

## TME adjustment specification grid

```{r}
show_head("results/tables/tme_adjustment_specification_summary.tsv", 20)
```

## Tumor-associated retained under S1 stromal/immune-only adjustment

```{r}
show_head("results/tables/program_tumor_intrinsic_retained_signature.tsv", 40)
```

## Composition-covarying signature

```{r}
show_head("results/tables/program_composition_covarying_signature.tsv", 40)
```

## Transition module external validation

```{r}
show_head("results/tables/transition_module_external_validation_summary.tsv", 20)
```

```{r}
show_head("results/tables/transition_module_external_validation.tsv", 100)
```

# Hub prioritization

## Original integrated hub table

```{r}
show_head("results/tables/integrative_hub_gene_prioritization.tsv", 40)
```

## Strict hub prioritization for manuscript triage

```{r}
show_head("results/tables/integrative_hub_gene_prioritization_strict.tsv", 50)
```

## Main-text hub candidates

```{r}
show_head("results/tables/main_text_hub_candidates.tsv", 50)
```

## Druggability annotation

```{r}
show_head("results/tables/druggability_annotation.tsv", 80)
```

# Key figures

```{r, results='asis'}
figs <- list.files("results/figures", pattern = "\\.png$", full.names = TRUE)
for (f in figs) {
  cat("\n\n## ", basename(f), "\n\n", sep = "")
  cat("![](", normalizePath(f, winslash = "/", mustWork = FALSE), ")\n\n", sep = "")
}
```
