---
title: "External validation of the cross-cohort PDAC transcriptomic atlas"
subtitle: "v3.14 composition-identifiability audit, external reinforcement, scRNA mapping, and conservative annotation"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
root <- normalizePath(".", mustWork = TRUE)
read_tsv_if <- function(path) if (file.exists(path)) readr::read_tsv(path, show_col_types = FALSE) else tibble::tibble(status = "missing", file = path)
knitr::opts_knit$set(root.dir = root)
```

# Purpose

This report summarizes the external reinforcement analyses after the v3.14 methodological correction. The main interpretation is now composition-identifiability rather than target discovery: tumor-versus-control bulk signatures are evaluated for reproducibility, TCGA/GTEx reinforcement, survival association, single-cell compartment localization, within-tumor deconvolution covariation, and conservative DepMap/Open Targets annotation.

# v3.14 composition-identifiability audit

```{r tme-specs}
specs <- read_tsv_if("results/tables/tme_adjustment_specifications.tsv")
knitr::kable(specs)
```

```{r tme-spec-summary}
sumtab <- read_tsv_if("results/tables/tme_adjustment_specification_summary.tsv")
knitr::kable(sumtab, digits = 4)
```

```{r deconv-cor-audit}
cora <- read_tsv_if("results/tables/gene_deconvolution_correlation_group_audit.tsv")
knitr::kable(cora, digits = 4)
```

```{r transition-id-audit}
tria <- read_tsv_if("results/tables/transition_program_identifier_summary.tsv")
knitr::kable(tria, digits = 4)
```

# TCGA-PAAD / GTEx pancreas validation

```{r tcga-gtex}
tcga <- read_tsv_if("results/tables/external/tcga_gtex_module_validation.tsv")
knitr::kable(tcga, digits = 4)
```

```{r tcga-fig, out.width='100%'}
fig <- "results/figures/external/tcga_gtex_module_scores.png"
if (file.exists(fig)) knitr::include_graphics(fig)
```

```{r tcga-signature-overlap}
ov <- read_tsv_if("results/tables/external/tcga_gtex_signature_gene_overlap.tsv")
knitr::kable(ov, digits = 4)
```

# TCGA survival association

```{r tcga-survival}
surv <- read_tsv_if("results/tables/external/tcga_survival_results.tsv")
knitr::kable(surv, digits = 4)
```

# Single-cell / single-nucleus mapping

```{r scrna-score}
sc <- read_tsv_if("results/tables/external/scrna_celltype_module_score_summary.tsv")
celltype_order <- c(
  "Epithelial (malignant)", "Epithelial (non-malignant)",
  "Cancer-associated fibroblast", "Myeloid", "Lymphoid",
  "Endothelial", "Endocrine", "Pericyte",
  "Vascular smooth muscle", "Schwann", "Adipocyte"
)
if (all(c("cell_type", "module") %in% names(sc))) {
  sc$cell_type <- factor(sc$cell_type, levels = celltype_order)
  sc <- sc[order(sc$module, is.na(sc$cell_type), sc$cell_type), , drop = FALSE]
}
knitr::kable(sc, digits = 4)
```

```{r scrna-fig, out.width='100%'}
fig <- "results/figures/external/scrna_module_celltype_heatmap.png"
if (file.exists(fig)) knitr::include_graphics(fig)
```

```{r scrna-signature-overlap}
ov <- read_tsv_if("results/tables/external/scrna_signature_gene_overlap.tsv")
knitr::kable(ov, digits = 4)
```

```{r scrna-dot}
dot <- read_tsv_if("results/tables/external/scrna_hub_gene_dotplot_table.tsv")
key_groups <- c(
  "Epithelial (malignant)", "Epithelial (non-malignant)",
  "Cancer-associated fibroblast", "Myeloid", "Lymphoid",
  "Endothelial", "Endocrine", "Pericyte"
)
key_genes <- c("SLC6A14", "LY75", "FGD6", "COL12A1", "ASAP2", "KIF18A", "RUVBL1", "PRIM2", "CTNNB1", "TUBB")
if (all(c("group", "gene") %in% names(dot))) {
  dot$key_group <- dot$group %in% key_groups
  dot$key_gene <- dot$gene %in% key_genes
  dot$group <- factor(dot$group, levels = key_groups)
  dot_show <- dot[order(!dot$key_group, is.na(dot$group), dot$group, !dot$key_gene, dot$gene), , drop = FALSE]
  dot_show$key_group <- NULL
  dot_show$key_gene <- NULL
} else {
  dot_show <- dot
}
knitr::kable(utils::head(dot_show, 120), digits = 4)
```

```{r scrna-dot-fig, out.width='100%'}
fig <- "results/figures/external/scrna_hub_gene_dotplot.png"
if (file.exists(fig)) knitr::include_graphics(fig)
```

# Deconvolution sensitivity

```{r deconv-effects}
dec <- read_tsv_if("results/tables/external/deconv_sensitivity_condition_effects.tsv")
knitr::kable(utils::head(dec, 80), digits = 4)
```

```{r deconv-fig, out.width='100%'}
fig <- "results/figures/external/deconv_sensitivity_condition_effects.png"
if (file.exists(fig)) knitr::include_graphics(fig)
```

```{r deconv-overlap}
ov <- read_tsv_if("results/tables/external/deconv_sensitivity_tme_class_overlap.tsv")
knitr::kable(ov, digits = 4)
```

# DepMap / Open Targets target prioritization

```{r target-pri}
tar <- read_tsv_if("results/tables/external/external_target_prioritization.tsv")
if ("main_text_candidate_score" %in% names(tar)) {
  tar <- tar[order(-tar$main_text_candidate_score, -tar$external_target_priority_score), , drop = FALSE]
}
target_cols <- intersect(c(
  "gene", "candidate_lane", "main_text_candidate_score", "external_target_priority_score",
  "tme_class", "tme_class_legacy", "classification_note", "retained_after_tme_adjustment", "meta_logFC", "adjusted_meta_logFC", "adjusted_fdr",
  "target_class", "tractability_manual", "subcellular_location",
  "depmap_expression_status", "pancreas_expression_percentile",
  "depmap_crispr_status", "median_gene_effect_pancreas", "pancreas_dependency_fraction_lt_minus05",
  "opentargets_status", "pancreas_disease_name", "pancreas_disease_score", "pancreas_known_drug_count"
), names(tar))
knitr::kable(utils::head(tar[, target_cols, drop = FALSE], 80), digits = 4)
```

```{r target-fig, out.width='100%'}
fig <- "results/figures/external/external_target_prioritization.png"
if (file.exists(fig)) knitr::include_graphics(fig)
```

# Output manifest

```{r manifest}
manifest <- read_tsv_if("results/tables/external/external_validation_output_manifest.tsv")
knitr::kable(manifest)
```
